---
title: "Presenting Regression Results Using R"
subtitle: "Best Practices for Statistical Communication"

author: |
  Mohammed Sohail
  Graduate School of Business, Hochschule Fresenius
  Matriculation No. 401080440
  sohail7853@gmail.com

date: "`r format(Sys.Date(), '%B %d, %Y')`"

format:
  revealjs:
    theme: simple
    slide-number: true
    progress: true
    chalkboard: false
    footer: "Presenting Regression Results Using R | Mohammed Sohail"
    transition: slide
    highlight-style: github
    code-fold: false
    fig-align: center

bibliography: references.bib
csl: apa.csl
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(stargazer)
library(broom)
library(knitr)

data(mtcars)
model1 <- lm(mpg ~ wt + cyl, data = mtcars)
model2 <- lm(mpg ~ wt + cyl + hp, data = mtcars)
```

# Presenting Regression Results Using R {background-color="#2C3E50"}

## Session Overview

By the end of this session, you will be able to:

::: incremental
- Interpret regression output from R correctly
- Produce **professional tables** using `stargazer`
- Create **coefficient plots** with `ggplot2`
- Build **prediction tables** from model output
- Follow **APA 7th edition** reporting conventions
:::

::: {.notes}
Welcome everyone. Today we focus not just on running regression but on presenting results in a way that is credible, clear, and follows disciplinary standards.
:::

# Part 1: What is Regression? {background-color="#34495E"}

## The Big Question

Regression answers: **How does X affect Y?**

| Research Question | Outcome (Y) | Predictor (X) |
|---|---|---|
| Does weight affect fuel efficiency? | MPG | Weight (lbs) |
| Does study time affect exam score? | Score | Hours studied |
| Does price affect sales volume? | Units sold | Price |

::: {.notes}
Give a real-world example from the students' field of study. The mtcars dataset is our running example throughout.
:::

## The Regression Equation

$$Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \epsilon$$

- $Y$ = **outcome** (what we want to predict)
- $\beta_0$ = **intercept** (value of Y when all X = 0)
- $\beta_1, \beta_2$ = **slope coefficients** (effect of each X on Y)
- $\epsilon$ = **error** (unexplained variation)

> The coefficients are the story. Everything else supports them.

## What Statistics to Report (APA Style)

| Statistic | What it tells you | APA notation |
|---|---|---|
| Coefficient ($\hat{\beta}$) | Direction and size of effect | $\hat{\beta}$ = -3.19 |
| Standard Error | Precision of estimate | *SE* = 0.74 |
| t-statistic | Signal-to-noise ratio | *t*(29) = -4.32 |
| p-value | Probability under null | *p* < .001 |
| 95% CI | Range of plausible values | [-4.69, -1.69] |
| R² | Total variance explained | *R*² = .83 |

# Part 2: Running the Model {background-color="#34495E"}

## The Code (3 Steps)

```{r}
#| echo: true
#| eval: false

# Step 1: Load data
data(mtcars)

# Step 2: Fit the regression model
model1 <- lm(mpg ~ wt + cyl, data = mtcars)

# Step 3: View results
summary(model1)
```

## Reading the Output

```{r}
#| echo: false
#| message: false

summary(model1)$coefficients |>
  as.data.frame() |>
  round(3) |>
  knitr::kable(col.names = c("Estimate", "Std. Error", "t value", "Pr(>|t|)"))
```

::: {.callout-tip}
**Key numbers**: Estimate = effect size | Pr(>|t|) = p-value | Look for p < .05
:::

# Part 3: Professional Tables {background-color="#34495E"}

## Why Not Just Use `summary()`?

::: columns
::: {.column width="50%"}
**summary() output**
- Designed for diagnostics
- Not publication-ready
- No standard error clarity
- Hard to compare models
:::

::: {.column width="50%"}
**stargazer output**
- APA-compatible layout
- Side-by-side model comparison
- Clear notation
- Export to LaTeX / text
:::
:::

## Building a Stargazer Table

```{r}
#| echo: true
#| message: false
#| results: asis

stargazer(
  model1, model2,
  type = "text",
  dep.var.labels = "Miles per Gallon",
  covariate.labels = c("Weight (1000 lbs)", "Cylinders", "Horsepower"),
  column.labels = c("Model 1", "Model 2"),
  keep.stat = c("n", "rsq", "adj.rsq", "f"),
  star.cutoffs = c(.05, .01, .001)
)
```

## Interpreting the Table

::: incremental
- **Weight** ($\hat{\beta}$ = -3.19, *p* < .001): Each additional 1,000 lbs reduces fuel efficiency by 3.19 mpg
- **Cylinders** ($\hat{\beta}$ = -1.51, *p* < .001): Each additional cylinder reduces efficiency by 1.51 mpg
- **R² = .83**: The model explains 83% of variation in fuel efficiency
- **Model 2 vs. Model 1**: Adding Horsepower provides minimal improvement ($\Delta R^2$ = .01)
:::

# Part 4: Coefficient Plots {background-color="#34495E"}

## Why Visualize Coefficients?

> "A coefficient plot shows what a regression table shows, but faster."

- Audiences can immediately see **direction** and **magnitude**
- Confidence intervals become visual — easy to judge significance
- Works for papers, reports, and presentations

## The Plot

```{r}
#| echo: true
#| message: false
#| fig-height: 3.5

results <- tidy(model1, conf.int = TRUE) %>%
  filter(term != "(Intercept)") %>%
  mutate(term = recode(term, "wt" = "Weight (1000 lbs)", "cyl" = "Cylinders"))

ggplot(results, aes(y = term, x = estimate)) +
  geom_point(size = 3, color = "#2C3E50") +
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high), width = 0.2, color = "#2C3E50") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "#E74C3C") +
  labs(x = "Coefficient Estimate", y = NULL,
       caption = "Error bars = 95% CI. Dashed line = zero effect.") +
  theme_minimal(base_size = 14)
```

# Part 5: Prediction Tables {background-color="#34495E"}

## Translating Results to Decisions

Prediction tables answer: **"What does the model predict for specific scenarios?"**

Useful for:
- Business presentations
- Policy recommendations
- Communicating to non-statisticians

## Building a Prediction Table

```{r}
#| echo: true
#| message: false
#| html-math-method: plain

scenarios <- tibble(wt = c(2.0, 2.5, 3.0, 3.5, 4.0), cyl = 6)
pred_out  <- predict(model1, newdata = scenarios, se.fit = TRUE)

scenarios %>%
  mutate(
    `Predicted MPG` = round(pred_out$fit, 1),
    `95% CI`        = paste0(
      "[", round(pred_out$fit - 1.96 * pred_out$se.fit, 1),
      ", ", round(pred_out$fit + 1.96 * pred_out$se.fit, 1), "]"
    )
  ) %>%
  rename(`Weight (1000 lbs)` = wt, Cylinders = cyl) %>%
  knitr::kable()
```

# Part 6: APA Reporting {background-color="#34495E"}

## Writing Up Results in Text

> A multiple linear regression was conducted predicting fuel efficiency from vehicle weight and cylinder count. The model was significant, *F*(2, 29) = 69.03, *p* < .001, *R*² = .83. Vehicle weight was a significant negative predictor ($\hat{\beta}$ = -3.19, *SE* = 0.74, *t*(29) = -4.32, *p* < .001, 95% CI [-4.69, -1.69]).

## APA Formatting Quick Reference

| Do this | Not this |
|---|---|
| *p* = .041 | p < 0.05 |
| *R*² = .83 | R2 = 0.83 |
| 95% CI [2.1, 4.3] | CI: 2.1-4.3 |
| *F*(2, 29) = 69.03 | F = 69 |
| *M* = 20.09, *SD* = 6.03 | Mean = 20 |

# Part 7: Check Assumptions {background-color="#34495E"}

## Before You Report: Diagnostics

```{r}
#| echo: true
#| fig-height: 3.5

par(mfrow = c(1, 2))
plot(model1, which = c(1, 2))
par(mfrow = c(1, 1))
```

## What to Check

| Assumption | Plot | What to look for |
|---|---|---|
| Linearity | Residuals vs. Fitted | No pattern — flat line |
| Homoscedasticity | Residuals vs. Fitted | Consistent spread |
| Normality | Q-Q Plot | Points on diagonal |
| Independence | Study design | Random sample |

# Student Exercise {background-color="#1A252F"}

## Your Turn (10 Minutes)

```{r}
#| echo: true
#| eval: false

# Load data
data(mtcars)

# Task 1: Fit a regression — predict mpg from horsepower only
model_hp <- lm(mpg ~ hp, data = mtcars)

# Task 2: Get summary
summary(model_hp)

# Task 3: Make a coefficient plot using broom + ggplot2
library(broom)
results_hp <- tidy(model_hp, conf.int = TRUE)

# Task 4: Write one APA-style sentence describing the hp coefficient
# Hint: "Horsepower was a significant predictor of fuel efficiency..."
```

::: {.callout-note}
See the **handout** for worked examples, APA table templates, and diagnostic plots.
:::

# Wrap-Up {background-color="#2C3E50"}

## Key Takeaways

::: incremental
1. **Tables** (stargazer) — for academic/formal reporting
2. **Coefficient plots** — for presentations and general audiences
3. **Prediction tables** — for business decisions and stakeholders
4. **Check assumptions** — before reporting anything
5. **APA style** — exact p-values, italics, CIs always
6. **Context** — numbers do not speak for themselves
:::

## Resources

- See the **handout.qmd** for full APA write-up template
- @james2013 — Introduction to Statistical Learning
- @wickham2019 — ggplot2 / tidyverse reference
- @fox2011 — R Companion to Applied Regression
- @r2024 — R language documentation

## Thank You! {background-color="#2C3E50"}

**Questions?**

---

*Slides and handout available in the course repository.*

# References {.unnumbered}

# Test workflow trigger - GitHub Actions will auto-render this file

::: {#refs}
:::
