---
title: "Presenting Regression Results Using R"
author: "Mohammed"
date: "`r format(Sys.time(), '%d %B %Y')`"
format: revealjs
---

# Presenting Regression Results Using R

## Today's Goal

Learn how to present regression results professionally using R

**Why this matters:**
- Running a regression ≠ presenting it well
- Good presentation = better understanding
- Different formats for different audiences

# What is Regression?

## The Big Picture

Regression answers: **How does X affect Y?**

Example:
- Does **weight** affect **fuel efficiency**?
- Does **experience** affect **salary**?
- Does **price** affect **sales**?

## The Simple Equation

$$Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \epsilon$$

- $Y$ = outcome (what we want to predict)
- $X$ = predictor (what influences the outcome)
- $\beta$ = effect size (how much X affects Y)

# Regression in 60 Seconds

## The Code

```{r}
#| echo: true
#| eval: false

# Step 1: Load data
data(mtcars)

# Step 2: Fit model
model <- lm(mpg ~ wt + cyl, data = mtcars)

# Step 3: See results
summary(model)
```

## What the Results Mean

- **Coefficient for `wt`** = -3.2: Each additional 1000 lbs → -3.2 mpg (fuel gets worse)
- **P-value** = 0.0001: This effect is real (not by chance)
- **R² = 0.83**: The model explains 83% of fuel efficiency variation

# Presenting Results: Option 1 - Tables

## Publication Table

```{r}
#| echo: true
#| message: false

library(stargazer)
library(tidyverse)

data(mtcars)
model1 <- lm(mpg ~ wt + cyl, data = mtcars)
model2 <- lm(mpg ~ wt + cyl + hp, data = mtcars)

stargazer(model1, model2, type = "text")
```

**When to use:** Academic papers, formal reports

# Presenting Results: Option 2 - Charts

## Coefficient Plot

```{r}
#| echo: true
#| message: false

library(broom)

results <- tidy(model1, conf.int = TRUE)

ggplot(results %>% filter(term != "(Intercept)"),
       aes(y = term, x = estimate)) +
  geom_point(size = 3, color = "darkblue") +
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high),
                width = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Effect Sizes with 95% Confidence Intervals",
       x = "Coefficient", y = "") +
  theme_minimal()
```

**When to use:** Presentations to non-technical audiences

# Presenting Results: Option 3 - Predictions

## Show Real-World Impact

```{r}
#| echo: true
#| message: false

scenarios <- tibble(wt = c(2.5, 3, 3.5, 4), cyl = 6)
pred <- predict(model1, newdata = scenarios, se.fit = TRUE)

scenarios %>%
  mutate(
    mpg = round(pred$fit, 1),
    ci = paste0("[", round(pred$fit - 1.96*pred$se.fit, 1),
                ", ", round(pred$fit + 1.96*pred$se.fit, 1), "]")
  ) %>%
  select(wt, cyl, mpg, ci) %>%
  knitr::kable(
    col.names = c("Weight (1000 lbs)", "Cylinders", "Predicted MPG", "95% CI")
  )
```

**When to use:** Business presentations, stakeholders

# Best Practices

## 1. Know Your Audience

| Audience | Format | Focus |
|----------|--------|-------|
| Statisticians | Tables, detailed statistics | Coefficients, p-values, R² |
| Business | Charts, scenarios | Real-world impact, predictions |
| General public | Simple visuals | "What does this mean for me?" |

## 2. Always Show Uncertainty

Don't just say: "The effect is 3.2"

Say: "The effect is 3.2 (95% CI: 2.1 to 4.3)"

## 3. Validate Assumptions

Before presenting, check:
- Linearity (is the relationship linear?)
- Independence (are observations independent?)
- Normality (are residuals normally distributed?)

## 4. Interpret in Context

**Bad:** "The coefficient is -3.2, p = 0.0001"

**Good:** "Each additional 1000 lbs of weight reduces fuel efficiency by 3.2 miles per gallon, a highly significant effect"

## 5. Avoid Common Mistakes

- ❌ Correlation = Causation
- ❌ Forgetting to mention sample size
- ❌ Ignoring outliers
- ❌ Over-interpreting small effects

# Your Turn: Student Exercise

## Exercise (10 minutes)

```{r}
#| echo: true
#| eval: false

# Load data
data(mtcars)

# Task 1: Fit a regression model
# Predict: mpg ~ hp (horsepower)
model <- lm(mpg ~ hp, data = mtcars)

# Task 2: Get summary
summary(model)

# Task 3: Interpret
# Q1: Does horsepower affect fuel efficiency?
# Q2: What is the effect size?
# Q3: Is it statistically significant?
```

**You have 10 minutes. Go!**

# Wrap-Up

## Key Takeaways

1. **Regression** = understand relationships between variables
2. **Tables** = detailed, formal
3. **Charts** = intuitive, accessible
4. **Predictions** = real-world impact
5. **Context** = numbers don't speak for themselves

---

## Thank You!

**Questions?**
